---
title: "Chapter 5"
author: "Wal McConnell"
date: "11 July 2017"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```


## 5.2 Building a Plot

Building a plot manually using layer function

```{r}
p <- ggplot(mpg, aes(displ, hwy))
p + layer(
  mapping = NULL,
  data = NULL,
  geom = "point", 
  stat = "identity", 
  position = "identity"
)
```

```{r}
model <- loess(hwy ~ displ, data = mpg)
grid <- data_frame(displ = seq(min(mpg$displ), max(mpg$displ), length = 50))
grid$hwy <- predict(model, newdata = grid)
std_resid <- resid(model) / model$s
outlier <- filter(mpg, abs(std_resid) > 2)
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_line(data = grid, colour = "blue", size = 1.5) +
  geom_text(data = outlier, aes(label = model))
```

## Exercises 5.3.1

### Question 2

```{r}
class <- mpg %>%
     group_by(class) %>%
     summarise(n = n(), hwy = mean(hwy))

ggplot(class, aes(x = class, y = hwy)) +
  geom_point(data = mpg, position = position_jitter(0.25)) +
  geom_point(color = "red", size = 5) +
  geom_text(aes(y = 10, label = paste("n = ",n)))
  
```


## 5.4 Aesthetic mappings

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme(legend.position = "none")


ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(method = "lm", se = FALSE) +
  theme(legend.position = "none")
```

```{r}
ggplot(mpg, aes(cty, hwy)) + geom_point(colour = "darkblue")

ggplot(mpg, aes(cty, hwy)) + geom_point(aes(colour = "darkblue"))
```

### Exercise 5.4.3

```{r}
ggplot(mpg) + geom_point(aes(mpg$displ, mpg$hwy))
ggplot(mpg, aes(displ, hwy)) + geom_point()
```
```{r}
ggplot() +
    geom_point(mapping = aes(y = hwy, x = cty), data = mpg) +
    geom_smooth(data = mpg, mapping = aes(cty, hwy))

ggplot(mpg, aes(cty, hwy)) + 
         geom_point() +
         geom_smooth()

```

```{r}
ggplot(diamonds, aes(carat, price)) +
     geom_point(aes(log(brainwt), log(bodywt)), data = msleep)

ggplot(msleep) +
     geom_point(aes(log(brainwt), log(bodywt)))

```

### Question 2

```{r}
ggplot(mpg) +
     geom_point(aes(class, cty)) +
     geom_boxplot(aes(trans, hwy))


ggplot(mpg) +
     geom_point(aes(class, cty))

ggplot(mpg) +
     geom_boxplot(aes(trans, hwy))

```
 ### Question 3
 
 
 
```{r}
library(precis)
precis(diamonds)

ggplot(diamonds, aes(y = color)) +
  geom_point(aes(x = clarity), color = "red") +
  geom_point(aes(x = carat), color = "blue")
  
```

### Exercises 5.5.1

| geom          | Aesthetics           
| ------------- |:-------------:
|geom_blank     | none
|geom_point     | x, y, alpha, colour, fill, group, shape, size, stroke
|geom_path      | x, y, alpha, colour, group, linetype, size
|geom_ribbon    | x, ymin, ymax, alpha, colour, fill, linetype, size
|geom_segment   | x, y, xend, yend, alpha, colour, group, linetype, size
|geom_rect      | x, y, alpha, colour, fill, linetype, size
|geom_polygon   | x, y, alpha, colour, fill, linetype, size
|geom_text      | x, y, label, alpha, angle, colour, family, fontface, group, hjust, lineheight, size, vjust

### Question 4

```{r}
ggplot(mpg, aes(drv, displ)) + 
  geom_violin()
```

```{r}
ggplot(mpg, aes(hwy, cty)) + 
  geom_count(alpha = 1/3) + 
  theme(legend.position = "none")

# This is a variant geom_point that counts the number of observations at each location, then maps the count to point area. It useful when you have discrete data and overplotting.

ggplot(mpg, aes(hwy, cty)) + 
  geom_point(alpha = 1/3) + 
  theme(legend.position = "none")
```



```{r}
ggplot(mpg, aes(hwy, cty)) + 
  geom_hex(bins = 10) + 
  theme(legend.position = "none")
```

```{r}

# The jitter geom is a convenient shortcut for geom_point(position = "jitter"). It adds a small amount of random variation to the location of each point, and is a useful way of handling overplotting caused by discreteness in smaller datasets.

ggplot(mpg, aes(cyl, drv)) + 
  geom_jitter(width = 0.2, height = 0.2)
```

```{r}
ggplot(economics, aes(date, psavert)) + geom_area()
```

```{r}
ggplot(economics, aes(uempmed, psavert)) + geom_path()
```


| Problem     | geom
| ------------- |:-------------:
| Display how a variable has changed over time. | geom_line
| Show the detailed distribution of a single variable. | geom_boxplot
| Focus attention on the overall trend in a large dataset. | geom_histogram
| Draw a map. | geom_map
| Label outlying points. | geom_text


### Exercises 5.6.2

```{r}

ggplot(mpg, aes(hwy, displ)) + geom_smooth()

mod <- loess(hwy ~ displ, data = mpg)
smoothed <- data.frame(displ = seq(1.6, 7, length = 50))
pred <- predict(mod, newdata = smoothed, se = TRUE)
smoothed$hwy <- pred$fit
smoothed$hwy_lwr <- pred$fit - 1.96 * pred$se.fit
smoothed$hwy_upr <- pred$fit + 1.96 * pred$se.fit

ggplot(smoothed, aes(y = displ, x = hwy)) +
  geom_ribbon(aes(ymin = min(displ), ymax = max(displ)), fill = "grey70")
  
```


```{r}
ggplot(mpg, aes(displ)) + stat_ecdf()
ggplot(mpg, aes(sample = displ)) + stat_qq()


ggplot(mpg, aes(displ)) + 
  geom_density(color = "pink") +
  stat_function(fun = "dnorm", args = list(mean = mean(mpg$displ), sd = sd(mpg$displ)))
  #stat_function(fun = "dnorm")


ggplot(mpg, aes(displ)) + 
  geom_density(color = "pink") +
  #stat_function(fun = "dnorm", args = list(mean = mean(mpg$displ), sd = sd(mpg$displ)))
  stat_function(fun = "dnorm")


```
```{r}
ggplot(mpg, aes(drv, trans)) + geom_count(aes(size = ..prop.., group = 1))
ggplot(mpg, aes(drv, trans)) + geom_count(aes(size = ..prop..))
```

