---
title: "R Notebook"
output: html_notebook
---


There are around 40 unique elements that control the appearance of the plot. They can be roughly grouped into five categories: plot, axis, legend, panel and facet.


```{r}
base + theme(plot.background = element_rect(colour = "grey50", size = 2))
base + theme(
  plot.background = element_rect(colour = "grey50", size = 2),
  plot.margin = margin(2, 2, 2, 2)
)
base + theme(plot.background = element_rect(fill = "lightblue"))
```

```{r}
df <- data.frame(x = 1:3, y = 1:3)
base <- ggplot(df, aes(x, y)) + geom_point()

# Accentuate the axes
base + theme(axis.line = element_line(colour = "grey50", size = 1))
# Style both x and y axis labels
base + theme(axis.text = element_text(color = "blue", size = 12))
# Useful for long labels
base + theme(axis.text.x = element_text(angle = -90, vjust = 0.5))
```

```{r}
df <- data.frame(
  x = c("label", "a long label", "an even longer label"), 
  y = 1:3
)
base <- ggplot(df, aes(x, y)) + geom_point()
base
base + 
  theme(axis.text.x = element_text(angle = -30, vjust = 1, hjust = 0)) + 
  xlab(NULL) + 
  ylab(NULL)
```



## Legend

```{r}
df <- data.frame(x = 1:4, y = 1:4, z = rep(c("a", "b"), each = 2))
base <- ggplot(df, aes(x, y, colour = z)) + geom_point()

base + theme(
  legend.background = element_rect(
    fill = "lemonchiffon", 
    colour = "grey50", 
    size = 1
  )
)
base + theme(
  legend.key = element_rect(color = "grey50"),
  legend.key.width = unit(0.9, "cm"),
  legend.key.height = unit(0.75, "cm")
)
base + theme(
  legend.text = element_text(size = 15),
  legend.title = element_text(size = 15, face = "bold")
)
```

# Panel Elements

```{r}
base <- ggplot(df, aes(x, y)) + geom_point()
# Modify background
base + theme(panel.background = element_rect(fill = "lightblue"))

# Tweak major grid lines
base + theme(
  panel.grid.major = element_line(color = "gray60", size = 0.8)
)
# Just in one direction  
base + theme(
  panel.grid.major.x = element_line(color = "gray60", size = 0.8)
)
```

```{r}
base2 <- base + theme(plot.background = element_rect(colour = "grey50"))
# Wide screen
base2 + theme(aspect.ratio = 9 / 16)
# Long and skiny
base2 + theme(aspect.ratio = 2 / 1)
# Square
base2 + theme(aspect.ratio = 1)
```


# Faceting elements
```{r}
df <- data.frame(x = 1:4, y = 1:4, z = c("a", "a", "b", "b"))
base_f <- ggplot(df, aes(x, y)) + geom_point() + facet_wrap(~z)

base_f
base_f + theme(panel.margin = unit(0.5, "in"))
base_f + theme(
  strip.background = element_rect(fill = "grey20", color = "grey80", size = 1),
  strip.text = element_text(colour = "white")
)
```

# Exercises

Create the ugliest plot possible! (Contributed by Andrew D. Steen, University of Tennessee - Knoxville)

theme_dark() makes the inside of the plot dark, but not the outside. Change the plot background to black, and then update the text settings so you can still read the labels.

Make an elegant theme that uses "linen" as the background colour and a serif font for the text.

Systematically explore the effects of hjust when you have a multiline title. Why doesn't vjust do anything?


## Saving the Plot

```{r}
pdf("output.pdf", width = 6, height = 6)
ggplot(mpg, aes(displ, cty)) + geom_point()
dev.off()


ggplot(mpg, aes(displ, cty)) + geom_point()
ggsave("output.pdf")
```


