---
title: "Chapter 9"
output: html_notebook
---


```{r}
library("tidyverse")
library("lubridate")
library("knitr")

```

The principle behind tidy data is simple: storing your data in a consistent way makes it easier to work with it.


Tidy data is a mapping between the statistical structure of a data frame (variables and observations) and the physical structure (columns and rows).


Tidy data follows two main principles:
1. Variables go in columns.
2. Observations go in rows.


The principle behind tidy data is simple: storing your data in a consistent way makes it easier to work with it. 
?
Index Data and Cartesian Data

Either form could be tidy depending on what the values "A", "B", "C", "D" mean.

[NOTE] Interesting point about missing values, explicit in one, implicit in the other

```{r}
indexed <- data.frame(
  x = c("a", "b", "c", "d", "c"),
  y = c("A", "D", "A", "C", "B"),
  z = c(1, 5, 4, 9, 10)
) %>% arrange(x, y)
matrix <- indexed %>% spread(y, z)

indexed
matrix
```


### Gathering 

Translating Cartesian to Indexed


```{r}

```


### Spreading

Translating Indexed data to Cartesian


# Tidyr

gather() and spread() functions


```{r}
ec2
gather(ec2, key = year, value = unemp, `2006`:`2015`)

# Gather all columns except for Month
gather(ec2, key = year, value = unemp, -month)
```

```{r}
ec2

gather(ec2, year, rate, `2006`:`2015`, na.rm = TRUE) # 112 rows
gather(ec2, year, rate, `2006`:`2015`, convert = TRUE) # 120 rows
# convert = TRUE automatically converts years from character strings to numbers 
gather(ec2, year, rate, `2006`:`2015`, convert = TRUE, na.rm = TRUE) # 112 rows
``` 

```{r}
ec2
economics_2 <- gather(ec2, year, rate, `2006`:`2015`, convert = TRUE, na.rm = TRUE)
economics_2
ggplot(economics_2, aes(year + (month - 1) / 12, rate)) + geom_line()

# Group by Year
ggplot(economics_2, aes(month, rate, group = year)) + geom_line(aes(colour = year), size = 1)
```

# Spread

```{r}
weather <- dplyr::data_frame(
  day = rep(1:3, 2),
  obs = rep(c("temp", "rain"), each = 3),
  val = c(c(23, 22, 20), c(0, 0, 5))
)
weather
spread(weather, key = obs, value = val)
```



# Exercises

### How can you translate each of the initial example datasets into the other form?


```{r}
economics <- ggplot2::economics
economics
gather(economics, key="metric", value="value", -date)
```




### How can you convert back and forth between the economics and economics_long datasets built into ggplot2?

```{r}
summary(economics)
```

```{r}
summary(economics_long)
```

# Seperate and Unite

```{r}
trt <- dplyr::data_frame(
  var = paste0(rep(c("beg", "end"), each = 3), "_", rep(c("a", "b", "c"))),
  val = c(1, 4, 2, 10, 5, 11)
)
trt
trt_seperated <- separate(trt, var, c("time", "treatment"), "_")
unite(trt_seperated, "var", c("time", "treatment"), sep="_")
```



# Exercises 2

Install the EDAWR package from https://github.com/rstudio/EDAWR. Tidy the who dataset.

```{r}
```


Work through the demos included in the tidyr package (demo(package = "tidyr"))

```{r}

```



# Blood Pressure Case Study

```{r}
bpd <- readr::read_table(
"name age      start  week1  week2  week3
Anne  35 2014-03-27 100/80 100/75 120/90
 Ben  41 2014-03-09 110/65 100/65 135/70
Carl  33 2014-04-02 125/80   <NA>   <NA>
", na = "<NA>")
```

```{r}
bpd
```

Convert to Cartesian form

```{r bpd}
bpd_1 <- gather(bpd, week, bp, week1:week3)
bpd_1
```

Seperate the bp variable


```{r bpd_1}
bpd_2 <- separate(bpd_1, bp, c("sys", "dia"), "/")
bpd_2
```

Pull out week number 

```{r bpd_2}
bpd_3 <- extract(bpd_2, week, "week", "(\\d)", convert = TRUE)
bpd_4 <- dplyr::arrange(bpd_3, name, week)
bpd_4
```


```{r}
ggplot(bpd_4, aes(x = week)) + geom_line(aes(y=dia)) + geom_line(aes(y=sys)) + facet_wrap(~ name)
```



# Test Scores Case Study

```{r}
scores <- dplyr::data_frame(
  person = rep(c("Greg", "Sally", "Sue"), each = 2),
  time   = rep(c("pre", "post"), 3),
  test1  = round(rnorm(6, mean = 80, sd = 4), 0),
  test2  = round(jitter(test1, 15), 0)
)
scores
```



```{r}
scores_1 <- gather(scores, test, score, test1:test2)
scores_1
```


```{r}
scores_2 <- spread(scores_1, time, score)
scores_2
```


```{r}
scores_3 <- mutate(scores_2, diff = post - pre)
scores_3
```


Plot it

```{r}
ggplot(scores_3, aes(person, diff, color = test)) +
  geom_hline(size = 2, colour = "white", yintercept = 0) +
  geom_point() +
  geom_path(aes(group = person), colour = "grey50",  arrow = arrow(length = unit(0.25, "cm")))
```

