---
title: "Chapter 11 - Removing Trend"
output:
  html_notebook: default
---

```{r echo=FALSE, warning=FALSE}
library(tidyverse)
```

# Removing Trend

So far our analysis of the diamonds data has been plagued by the powerful relationship between size (carat) and price

It makes it very difficult to see the impact of cut, colour and clarity because higher quality diamonds tend to be smaller, and hence cheaper


```{r echo=FALSE}
summary(diamonds)
ggplot(diamonds, aes(carat, price)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, colour = "yellow")


ggplot(diamonds, aes(carat, price, color = cut)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, size = 2)


ggplot(diamonds, aes(carat, price, color = clarity)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, size = 2)


ggplot(diamonds, aes(carat, price, color = color)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, size = 2)
```
Price rises sharper for higher quality diamonds!


```{r}
ggplot(diamonds %>% filter(carat <= 2), aes(carat, price)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, colour = "yellow")
```

```{r}
model <- lm(price ~ carat, data = diamonds)
summary(model)
coef(summary(model))
summary(resid(mod))
ggplot(diamonds %>% mutate(rel_price = resid(mod)), aes(carat, rel_price)) + geom_point()
ggplot(diamonds %>% mutate(rel_price = resid(mod)), aes(carat, rel_price)) + geom_bin2d()
```



We can use a linear model to remove the effect of size on price. Instead of looking at the raw price, we can look at the relative price: how valuable is this diamond relative to the average diamond of the same size.


```{r echo=FALSE}
diamonds2 <- diamonds %>% 
  filter(carat <= 2) %>%
  mutate(
    lcarat = log2(carat),
    lprice = log2(price)
  )
diamonds2

ggplot(diamonds2, aes(lcarat, lprice)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, colour = "yellow")
```




```{r echo=FALSE}
mod <- lm(lprice ~ lcarat, data = diamonds2)
summary(mod)
coef(summary(mod))
```

### Residuals

residuals are the vertical distance between each point and the line of best fit.


A relative price of zero means that the diamond was at the average price; positive means that it’s means that it’s more expensive than expected (based on its size), and negative means that it’s cheaper than expected.


```{r}
diamonds2 <- diamonds2 %>% mutate(rel_price = resid(mod))
ggplot(diamonds2, aes(carat, rel_price)) + 
  geom_bin2d()
```
```{r}
xgrid <- seq(-2, 1, by = 1/3)
data.frame(logx = xgrid, x = round(2 ^ xgrid, 2))
```

```{r}
color_cut <- diamonds2 %>% 
  group_by(color, cut) %>%
  summarise(
    price = mean(price), 
    rel_price = mean(rel_price)
  )
color_cut
```

```{r}
ggplot(color_cut, aes(color, price)) + 
  geom_line(aes(group = cut), colour = "grey80") +
  geom_point(aes(colour = cut))
```


```{r}
ggplot(color_cut, aes(color, rel_price)) + 
  geom_line(aes(group = cut), colour = "grey80") +
  geom_point(aes(colour = cut))
```



## Exercises

What happens if you repeat the above analysis with all diamonds? (Not just all diamonds with two or fewer carats). 

```{r}
ggplot(diamonds, aes(carat, price)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm")

ldiamonds <- diamonds %>% mutate(lcarat = log(carat), lprice = log(price))

ggplot(ldiamonds, aes(lcarat, lprice)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm")

model <- lm(lprice ~ lcarat, data = ldiamonds)
coef(summary(model))

ldiamonds <- ldiamonds %>% mutate(rel_price = resid(model))
ggplot(ldiamonds, aes(carat, price)) + geom_bin2d()
ggplot(ldiamonds, aes(carat, rel_price)) + geom_bin2d()
ggplot(ldiamonds, aes(lcarat, rel_price)) + geom_bin2d()

```

## What does the strange geometry of log(carat) vs relative price represent?

The largest diamonds tend to be cheaper than average because they are typically of lower quality


Are Lower quality Diamonds are larger ? 


```{r}

```





```{r}
by_carat <- ldiamonds %>% group_by(lcarat) %>%
  summarise(price = mean(price),
            rel_price = mean(rel_price))

ggplot(by_carat, aes(lcarat, rel_price)) + geom_point()


by_carat_cut <- ldiamonds %>% group_by(lcarat, cut) %>%
  summarise(price = mean(price),
            rel_price = mean(rel_price))

ggplot(by_carat_cut, aes(lcarat, rel_price)) + geom_point() + geom_smooth(method = "lm")


by_carat_color <- ldiamonds %>% group_by(lcarat, color) %>%
  summarise(price = mean(price), rel_price = mean(rel_price))

ggplot(by_carat_color, aes(lcarat, rel_price)) + geom_point()  + geom_smooth(method = "lm")


by_carat_clarity <- ldiamonds %>% group_by(lcarat, clarity) %>%
  summarise(price = mean(price), rel_price = mean(rel_price))

ggplot(by_carat_clarity, aes(lcarat, rel_price)) + geom_point()  + geom_smooth(method = "lm")


```

## I made an unsupported assertion that lower-quality diamonds tend to be larger. Support my claim with a plot.

Yes for clarity and color, less so for cut

```{r}
ggplot(diamonds, aes(carat, fill = clarity)) + geom_histogram()
ggplot(diamonds, aes(log(carat), fill = clarity)) + geom_histogram()

ggplot(diamonds, aes(carat, fill = color)) + geom_histogram()
ggplot(diamonds, aes(log(carat), fill = color)) + geom_histogram()

ggplot(diamonds, aes(carat, fill = cut)) + geom_histogram()
ggplot(diamonds, aes(log(carat), fill = cut)) + geom_histogram()

ggplot(diamonds, aes(carat, clarity, fill = clarity)) + geom_bin2d()
ggplot(diamonds, aes(carat, cut, fill = cut)) + geom_bin2d()
ggplot(diamonds, aes(carat, color, fill = color)) + geom_bin2d()


ggplot(diamonds) +  geom_smooth(method = 'lm', aes(carat, color)) 
ggplot(diamonds) +  geom_smooth(method = 'lm', aes(carat, clarity))
ggplot(diamonds) +  geom_smooth(method = 'lm', aes(carat, cut))


ggplot(diamonds, aes(log(carat), color=cut)) + geom_density()
ggplot(diamonds, aes(log(carat), color=clarity)) + geom_density()
ggplot(diamonds, aes(log(carat), color=color)) + geom_density()


```
How do depth and table relate to the relative price?


# Depth - no linear relationship

```{r}
cor(diamonds$price, diamonds$depth)
cor(log(diamonds$price), diamonds$depth)

ggplot(diamonds, aes(depth, price)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2)

ggplot(diamonds, aes(log(depth), log(price))) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2)

model <- lm(log(price) ~ depth, data = diamonds)
summary(model)

m_diamonds <- diamonds %>% mutate(rel_price = resid(model))
ggplot(m_diamonds, aes(depth, rel_price)) + geom_bin2d()

```

```{r}
cor(diamonds$price, diamonds$table)
cor(log(diamonds$price), diamonds$table)

ggplot(diamonds, aes(carat, price)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, color = "Yellow")

ldiamonds <- diamonds %>% mutate(lcarat = log(carat), lprice = log(price))

model <- lm(lprice ~ lcarat, data = ldiamonds)
coef(summary(model))

ldiamonds <- ldiamonds %>% mutate(rel_price = resid(model))

by_carat_depth <- ldiamonds %>% group_by(lcarat, depth) %>%
  summarise(price = mean(price), rel_price = mean(rel_price))

ggplot(by_carat_depth, aes(lcarat, rel_price)) + geom_point() + geom_smooth(method = "lm")


by_carat_table <- ldiamonds %>% group_by(lcarat, table) %>%
  summarise(price = mean(price), rel_price = mean(rel_price))

ggplot(by_carat_table, aes(lcarat, rel_price)) + geom_point() + geom_smooth(method = "lm")
```

