---
title: "Texas Housing Data"
output: html_notebook
---


```{r}
library(tidyverse)
```

# How sales have varied over time ? 

```{r}
ggplot(txhousing, aes(date, sales)) + geom_line(aes(group = city), alpha = 1/2)
```

## Plot Log scales to remove orders or magnitude

```{r}
ggplot(txhousing, aes(date, log(sales))) + geom_line(aes(group = city), alpha = 1/2)
```

## Use a categorical predictor to remove the month effect



By calculating the residual from the average for that month

### 1 city Abilene

```{r}
abilene <- txhousing %>% filter(city == "Abilene")
ggplot(abilene, aes(date, log(sales))) + geom_line()

mod <- lm(log(sales) ~ factor(month), data = abilene)
abilene$rel_sales <- resid(mod)
ggplot(abilene, aes(date, rel_sales)) + geom_line()
```

```{r}
txhousing <- txhousing %>%
  group_by(city) %>%
  mutate(rel_sales = resid(lm(log(sales) ~ factor(month), na.action = na.exclude)))


ggplot(txhousing, aes(date, rel_sales)) +
  geom_line(aes(group = city), alpha = 1/5) +
  geom_line(stat = "summary", fun.y = "mean", colour = "red")
```

## Exercises


The final plot shows a lot of short-term noise in the overall trend. How could you smooth this further to focus on long-term changes?

```{r}
txhousing <- txhousing %>% mutate(quarter = round(month / 4) + 1)

txhousing <- txhousing %>%
  group_by(city) %>%
  mutate(rel_sales = resid(lm(log(sales) ~ factor(quarter), na.action = na.exclude)))


ggplot(txhousing, aes(date, rel_sales)) +
  geom_line(aes(group = city), alpha = 1/5) +
  geom_line(stat = "summary", fun.y = "mean", colour = "red")
```

```{r}
txhousing <- txhousing %>%
  group_by(city) %>%
  mutate(rel_sales = resid(lm(log(sales) ~ factor(year), na.action = na.exclude)))


ggplot(txhousing, aes(date, rel_sales)) +
  geom_line(aes(group = city), alpha = 1/5) +
  geom_line(stat = "summary", fun.y = "mean", colour = "red")
```




```{r}
txhousing <- txhousing %>%
  group_by(city) %>%
  mutate(rel_sales = resid(lm(log(sales) ~ factor(month), na.action = na.exclude)))


ggplot(txhousing, aes(date, rel_sales)) +
  geom_line(aes(group = city), alpha = 1/5) +
  geom_line(stat = "summary", fun.y = "mean", colour = "red") + 
  xlim(2005, 2015)
```

# 2010 dip

```{r}
library(zoo)
txhousing_grouped_by_year_month <- txhousing %>% 
  group_by(year, month) %>% 
  summarise(sales = mean(sales, na.rm = TRUE)) %>%
  mutate(date = as.Date(as.yearmon(paste(year, month, sep="-"))))

ggplot(txhousing_grouped_by_year_month, aes(date, sales)) + geom_line(alpha = 1/5)
```


```{r}
data(txhousing)


ggplot(txhousing, aes(date, log(volume))) + geom_line(aes(group = city), alpha = 1/2)
ggplot(txhousing, aes(date, log(listings))) + geom_line(aes(group = city), alpha = 1/2)
ggplot(txhousing, aes(date, inventory)) + geom_line(aes(group = city), alpha = 1/2)

```
### Remove seasonality from volume

```{r}
ggplot(txhousing, aes(date, log(volume))) + geom_line(aes(group = city), alpha = 1/2)


txhousing <- txhousing %>%
  group_by(city) %>%
  mutate(rel_volume = resid(lm(log(volume) ~ factor(month), na.action = na.exclude)))


ggplot(txhousing, aes(date, rel_volume)) +
  geom_line(aes(group = city), alpha = 1/5) +
  geom_line(stat = "summary", fun.y = "mean", colour = "blue")
```

```{r}
ggplot(txhousing, aes(date, log(listings))) + geom_line(aes(group = city), alpha = 1/2)

txhousing <- txhousing %>%
  group_by(city) %>%
  mutate(rel_listings = resid(lm(log(listings) ~ factor(month), na.action = na.exclude)))


ggplot(txhousing, aes(date, rel_listings)) +
  geom_line(aes(group = city), alpha = 1/5) +
  geom_line(stat = "summary", fun.y = "mean", colour = "green")
```

### cities with incomplete time series

```{r}
data(txhousing)



incomplete_cases <- txhousing %>% filter(!complete.cases(.))

incomplete_cases_by_city <- incomplete_cases %>% group_by(city) %>% summarise(count = n()) %>% arrange(count)

ggplot(incomplete_cases_by_city, aes(city, count)) + 
  geom_bar(stat = "identity") + 
  coord_flip()

```

```{r}
?stat_summary

```

