---
title: "Chapter 11 - Removing Trend"
author: "Wal McConnell"
date: "9/3/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Removing Trend

* Analysis of diamonds dataset plagued by the powerful relationship between size and price. Difficult to see the impact of cut, color and clarity.


* Instead of looking at raw price, we look at relative price, how valuable that diamond is relative to the average diamond of the same size. 

This will help us see how much more expensive than average diamonds with better cut, color and clarity are.




## Diamonds Less than 2 carats

```{r}
ldiamonds <- diamonds %>% 
  filter(carat <= 2) %>%
  mutate(
    lcarat = log(carat),
    lprice = log(price)
  )

ldiamonds

ggplot(ldiamonds, aes(lcarat, lprice)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, colour = "yellow")

model <- lm(lprice ~ lcarat, data = ldiamonds)
coef(summary(model))

ldiamonds <- ldiamonds %>% mutate(rel_price = resid(model))

ggplot(ldiamonds, aes(lcarat, rel_price)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, colour = "yellow")

xgrid <- seq(-2, 1, by = 1/3)
data.frame(logx = xgrid, x = round(2 ^ xgrid, 2))

color_cut <- ldiamonds %>% 
  group_by(color, cut) %>%
  summarise(
    price = mean(price), 
    rel_price = mean(rel_price)
  )
color_cut

ggplot(color_cut, aes(color, price)) + 
  geom_line(aes(group = cut), colour = "grey80") +
  geom_point(aes(colour = cut))

ggplot(color_cut, aes(color, rel_price)) + 
  geom_line(aes(group = cut), colour = "grey80") +
  geom_point(aes(colour = cut))

ggplot(color_cut, aes(cut, rel_price)) + 
  geom_line(aes(group = color), colour = "grey80") +
  geom_point(aes(colour = color))
```

# Convert log relative price to actual multiples

```{r}
ggplot(ldiamonds, aes(lcarat, difference, color = cut)) + geom_point()
ggplot(ldiamonds, aes(lcarat, difference, color = color)) + geom_point()
ggplot(ldiamonds, aes(lcarat, difference, color = clarity)) + geom_point()
```

# Exercises

What happens if you repeat the above analysis with all diamonds? (Not just all diamonds with two or fewer carats). 

## All Diamonds

```{r}
ldiamonds <- diamonds %>% 
  mutate(
    lcarat = log(carat),
    lprice = log(price)
  )

ldiamonds

ggplot(ldiamonds, aes(lcarat, lprice)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, colour = "yellow")

model <- lm(lprice ~ lcarat, data = ldiamonds)
coef(summary(model))

ldiamonds <- ldiamonds %>% mutate(rel_price = resid(model))

ggplot(ldiamonds, aes(lcarat, rel_price)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, colour = "yellow")

```


## Only diamonds > 2 carats

```{r}
ldiamonds <- diamonds %>%
  filter(carat > 2) %>%
  mutate(
    lcarat = log(carat),
    lprice = log(price)
  )

ldiamonds

ggplot(ldiamonds, aes(lcarat, lprice)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, colour = "yellow")

model <- lm(lprice ~ lcarat, data = ldiamonds)
coef(summary(model))

ldiamonds <- ldiamonds %>% mutate(rel_price = resid(model))

ggplot(ldiamonds, aes(carat, rel_price)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, colour = "yellow")

```



What does the strange geometry of log(carat) vs relative price represent? What does the diagonal line without any points represent?

That the largest diamonds tend to be less expensive



# Question 2

I made an unsupported assertion that lower-quality diamonds tend to be larger. Support my claim with a plot.

```{r}
ggplot(diamonds, aes(carat, fill = clarity)) + geom_histogram()
ggplot(diamonds, aes(carat, fill = color)) + geom_histogram()
ggplot(diamonds, aes(carat, fill = cut)) + geom_histogram()

ggplot(diamonds, aes(carat, clarity, fill = clarity)) + geom_bin2d()
ggplot(diamonds, aes(carat, cut, fill = cut)) + geom_bin2d()
ggplot(diamonds, aes(carat, color, fill = color)) + geom_bin2d()


ggplot(diamonds) +  geom_smooth(method = 'lm', aes(carat, color)) 
ggplot(diamonds) +  geom_smooth(method = 'lm', aes(carat, clarity))
ggplot(diamonds) +  geom_smooth(method = 'lm', aes(carat, cut))


ggplot(diamonds, aes(log(carat), color=cut)) + geom_density()
ggplot(diamonds, aes(log(carat), color=clarity)) + geom_density()
ggplot(diamonds, aes(log(carat), color=color)) + geom_density()


```

How do depth and table relate to the relative price?


# Depth - no linear relationship

```{r}
cor(diamonds$price, diamonds$depth)
cor(log(diamonds$price), diamonds$depth)

ggplot(diamonds, aes(depth, price)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2)

ggplot(diamonds, aes(log(depth), log(price))) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2)

model <- lm(log(price) ~ depth, data = diamonds)
summary(model)

m_diamonds <- diamonds %>% mutate(rel_price = resid(model))
ggplot(m_diamonds, aes(depth, rel_price)) + geom_bin2d()

```

```{r}
cor(diamonds$price, diamonds$table)
cor(log(diamonds$price), diamonds$table)

ggplot(diamonds, aes(carat, price)) + 
  geom_bin2d() + 
  geom_smooth(method = "lm", se = FALSE, size = 2, color = "Yellow")

ldiamonds <- diamonds %>% mutate(lcarat = log(carat), lprice = log(price))

model <- lm(lprice ~ lcarat, data = ldiamonds)
coef(summary(model))

ldiamonds <- ldiamonds %>% mutate(rel_price = resid(model))

by_carat_depth <- ldiamonds %>% group_by(lcarat, depth) %>%
  summarise(price = mean(price), rel_price = mean(rel_price))

ggplot(by_carat_depth, aes(lcarat, rel_price)) + geom_point() + geom_smooth(method = "lm")


by_carat_table <- ldiamonds %>% group_by(lcarat, table) %>%
  summarise(price = mean(price), rel_price = mean(rel_price))

ggplot(by_carat_table, aes(lcarat, rel_price)) + geom_point() + geom_smooth(method = "lm")
```

