---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(ggplot2)
```

# Positioning - Faceting


Facetting generates small multiples each showing a different subset of the data


There are three types of facetting:

* facet_null(): a single plot, the default. \indexf{facet_null}
* facet_wrap(): "wraps" a 1d ribbon of panels into 2d.
* facet_grid(): produces a 2d grid of panels defined by variables which form the rows and columns.

```{r}
# Load Data
mpg2 <- subset(mpg, cyl != 5 & drv %in% c("4", "f") & class != "2seater")

base <- ggplot(mpg2, aes(displ, hwy)) + 
  geom_blank() + 
  xlab(NULL) + 
  ylab(NULL)

```

### Facet Wrap

as.table controls whether the facets are laid out like a table (TRUE), with highest values at the bottom-right, or a plot (FALSE), with the highest values at the top-right. dir controls the direction of wrap: horizontal or vertical.

### Why ?

```{r}

base + facet_wrap(~class, ncol = 3)
base + facet_wrap(~class, ncol = 3, as.table = FALSE)

base + facet_wrap(~class, nrow = 3)
base + facet_wrap(~class, nrow = 3, dir = "v")
```


### Facet Grid

```{r}
base + facet_grid(. ~ cyl)
base + facet_grid(drv ~ .)
base + facet_grid(drv ~ cyl)
```

### Controlling Scales


```{r}
p <- ggplot(mpg2, aes(cty, hwy)) + geom_jitter(width = 0.1, height = 0.1)
p + facet_wrap(~cyl)
p + facet_wrap(~cyl, scales = "free")


```

```{r}
economics_long
ggplot(economics_long, aes(date, value)) + 
  geom_line() + 
  facet_wrap(~variable, scales = "free_y", ncol = 1)
```
## [NOTE] reorder function

```{r}
mpg2$model <- reorder(mpg2$model, mpg2$cty)
mpg2$manufacturer <- reorder(mpg2$manufacturer, -mpg2$cty)
ggplot(mpg2, aes(cty, model)) + geom_point()
ggplot(mpg2, aes(cty, model)) + geom_point() + facet_grid(manufacturer ~ ., scales = "free", space = "free")
ggplot(mpg2, aes(cty, model)) + 
  geom_point() + 
  facet_grid(manufacturer ~ ., scales = "free", space = "free") +
  theme(strip.text.y = element_text(angle = 0)) # Rotates the manufacturer
```


### Missing faceting variables

Here ggplot will do what you expect: it will display the map in every facet: missing facetting variables are treated like they have all values.

```{r}
df1 <- data.frame(x = 1:3, y = 1:3, gender = c("f", "f", "m"))
df2 <- data.frame(x = 2, y = 2)

ggplot(df1, aes(x, y)) + 
  geom_point(data = df2, colour = "red", size = 2) + 
  geom_point() + 
  facet_wrap(~gender)
```

### Grouping v Faceting


#### [Note]- Putting all data in backgrond

```{r}
df <- data.frame(
  x = rnorm(120, c(0, 2, 4)),
  y = rnorm(120, c(1, 2, 1)),
  z = letters[1:3]
)

ggplot(df, aes(x, y)) + geom_point(aes(colour = z))
ggplot(df, aes(x, y)) + geom_point() + facet_wrap(~z)

df_sum <- df %>% 
  group_by(z) %>% 
  summarise(x = mean(x), y = mean(y)) %>%
  rename(z2 = z)
ggplot(df, aes(x, y)) + 
  geom_point() + 
  geom_point(data = df_sum, aes(colour = z2), size = 4) + 
  facet_wrap(~z)

df2 <- dplyr::select(df, -z)

ggplot(df, aes(x, y)) + 
  geom_point(data = df2, colour = "grey70") +
  geom_point(aes(colour = z)) + 
  facet_wrap(~z)
```

## Continuous Variables
```{r}
# Bins of width 1
mpg2$disp_w <- cut_width(mpg2$displ, 1)
# Six bins of equal length
mpg2$disp_i <- cut_interval(mpg2$displ, 6)
# Six bins containing equal numbers of points
mpg2$disp_n <- cut_number(mpg2$displ, 6)

plot <- ggplot(mpg2, aes(cty, hwy)) +
  geom_point() +
  labs(x = NULL, y = NULL)
plot + facet_wrap(~disp_w, nrow = 1)
plot + facet_wrap(~disp_i, nrow = 1)
plot + facet_wrap(~disp_n, nrow = 1)
```


## Exercises 7.2.7

#### Question 1
```{r}
diamonds
```

#### Question 2
```{r}
diamonds
```

#### Question 3
```{r}
ggplot(mpg2, aes(displ, hwy)) + 
  geom_smooth(data = select(mpg2, -class), se = FALSE) + 
  geom_point() + 
  facet_wrap(~class, nrow = 2)
```


