---
title: "Chapter 6 - Scales, Axes and Legends"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(precis)
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


Scales control the mapping from data to aesthetics.

Formally, each scale is a function from a region in data space (the domain of the scale) to a region in aesthetic space (the range of the scale). 


#### Scale Types
continuous position scales
colour scales
manual scales
identity scales.


```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = class))

ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = class)) +
  scale_x_continuous() + 
  scale_y_continuous() + 
  scale_colour_discrete()

```



```{r}
precis(mpg)
```


The use of + to "add" scales to a plot is a little misleading. When you + a scale, you're not actually adding it to the plot, but overriding the existing scale

```{r}
df <- data.frame(x = 1:2, y = 1, z = "a")
p <- ggplot(df, aes(x, y)) + geom_point()
p + scale_x_continuous("X axis")
p + scale_x_continuous(quote(a + mathematical ^ expression))
```


```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = class)) + 
  scale_x_continuous("A really awesome x axis label") +
  scale_y_continuous("An amazingly great y axis label")
```

 Using different scale
 
```{r}
base_plot <- ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = class))
base_plot + scale_x_sqrt()
base_plot + scale_x_continuous() 
base_plot + scale_x_discrete() # Why do the axis labels disappear if plotting a discrete variable onto a continuous scale ?
base_plot + scale_x_log10() # Why axis labels disappear ?
base_plot + scale_x_reverse()
base_plot + scale_x_time()
```

## Exercises 6.2.1

What happens if you pair a discrete variable to a continuous scale? What happens if you pair a continuous variable to a discrete scale?

```{r}
 precis(mpg)

mpg$year <- as.factor(mpg$year)

ggplot(mpg, aes(year, displ)) + 
  geom_point()

ggplot(mpg, aes(year, displ)) + 
  geom_point() +
  scale_y_discrete()
  
ggplot(mpg, aes(year, displ)) + 
  geom_point() +
  scale_x_continuous()

```

Simplify graph definitions 

```{r}

ggplot(mpg, aes(displ)) + 
  scale_y_continuous("Highway mpg") + 
  scale_x_continuous() +
  geom_point(aes(y = hwy))

is.factor(mpg$displ)
is.factor(mpg$hwy)

ggplot(mpg, aes(displ, hwy)) + 
  geom_point() +
  labs(y ="Highway mpg")

```

```{r}

is.factor(mpg$displ)
is.factor(mpg$class)

ggplot(mpg, aes(y = displ, x = class)) + 
  scale_y_continuous("Displacement (l)") + 
  scale_x_discrete("Car type") +
  scale_x_discrete("Type of car") + 
  scale_colour_discrete() + 
  geom_point(aes(colour = drv)) + 
  scale_colour_discrete("Drive\ntrain")

ggplot(mpg, aes(as.factor(class), displ, color = drv)) + 
  geom_point() + 
  labs(x = "Type of car", y = "Displacement (l)", color = "Drive\ntrain")
  
```

## Guides

If you're struggling to get the legend you want, it's likely that your data is in the wrong form

### Scale Title

```{r}
df <- data.frame(x = 1:2, y = 1, z = "a")

p <- ggplot(df, aes(x, y)) + geom_point()
p + scale_x_continuous("X axis")
p + scale_x_continuous(quote(a + mathematical ^ expression))

```

### Axis labels

```{r}
p <- ggplot(df, aes(x, y)) + geom_point(aes(colour = z))

p + xlab("X axis") + ylab("Y axis")

p + labs(x = "X axis", y = "Y axis", colour = "Colour\nlegend")
```

```{r}
p <- ggplot(df, aes(x, y)) + 
  geom_point() + 
  theme(plot.background = element_rect(colour = "grey50"))

p + labs(x = "x",  y = "y")
p + labs(x = NULL, y = NULL)
```


### Breaks and Legends

The breaks argument controls which values appear as tick marks on axes and keys on legends

```{r}
df <- data.frame(x = c(1, 3, 5) * 1000, y = 1)

axs <- ggplot(df, aes(x, y)) + 
  geom_point() + 
  labs(x = NULL, y = NULL)
axs
axs + scale_x_continuous(breaks = c(2000, 4000))
axs + scale_x_continuous(breaks = c(2000, 4000), labels = c("2k", "4k"))

# Show no breaks or scales
axs + scale_x_continuous(breaks = NULL) 
axs + scale_x_continuous(labels = NULL)

leg <- ggplot(df, aes(y, x, fill = x)) + 
  geom_tile() + 
  labs(x = NULL, y = NULL)
leg
leg + scale_fill_continuous(breaks = c(2000, 4000))
leg + scale_fill_continuous(breaks = c(2000, 4000), labels = c("2k", "4k"))
```

If you set labels, you must also set breaks

Rename breaks in categorical scale using labels

```{r}
df2 <- data.frame(x = 1:3, y = c("a", "b", "c"))

ggplot(df2, aes(x, y)) + 
  geom_point()

ggplot(df2, aes(x, y)) + 
  geom_point() + 
  scale_y_discrete(labels = c(a = "apple", b = "banana", c = "carrot"))
```

### Supply a function to breaks or labels

```{r}
df <- data.frame(x = c(1, 3, 5) * 1000, y = 1)

axs <- ggplot(df, aes(x, y)) + 
  geom_point() + 
  labs(x = NULL, y = NULL)

leg <- ggplot(df, aes(y, x, fill = x)) + 
  geom_tile() + 
  labs(x = NULL, y = NULL)

axs + scale_y_continuous(labels = scales::percent_format())
axs + scale_y_continuous(labels = scales::dollar_format("$"))
leg + scale_fill_continuous(labels = scales::unit_format("k", 1e-3)) # legend
leg + scale_y_continuous(labels = scales::unit_format("k", 1e-3)) # y axis

```

### Minor breaks

```{r}
df <- data.frame(x = c(2, 3, 5, 10, 200, 3000), y = 1)
ggplot(df, aes(x, y)) + 
  geom_point() + 
  scale_x_log10()

mb <- as.numeric(1:10 %o% 10 ^ (0:4))

ggplot(df, aes(x, y)) + 
  geom_point() + 
  scale_x_log10(minor_breaks = log10(mb)) # Shows more at lower end
```

## Exercises 6.3.3

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() +
  labs(x = "Displacement", y = quote(paste("Highway ", (frac("miles", "gallon"))))) +
  scale_x_continuous(labels = scales::unit_format("L"))
  
```

### Breaks v Labels

#### Breaks
* NULL for no breaks
* waiver() for the default breaks computed by the transformation object
* A numeric vector of positions
* A function that takes the limits as input and returns breaks as output

#### Labels

* NULL for no labels
* waiver() for the default labels computed by the transformation object
* A character vector giving labels (must be same length as breaks)
* A function that takes the breaks as input and returns labels as output

```{r}
data(mpg)
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = drv)) +
  scale_color_discrete(labels = c("4" = "4wd", "f" = "fwd", "r" = "rwd"))
```

### Question 4
What label function allows you to create mathematical expressions? What label function converts 1 to 1st, 2 to 2nd, and so on?

```{r}
library(scales)
ordinal(c(1,2,3,4,5,6))
```

### Question 5

What are the three most important arguments that apply to both axes and legends? What do they do? Compare and contrast their operation for axes vs. legends.


* breaks
* labels
* limits

```{r}

```



## Legends


A legend may need to draw symbols from multiple layers

```{r}
df <- data.frame(x = 1, y = 1:3, z = letters[1:3])
p <- ggplot(df, aes(x, y, colour = z))
p + geom_point()
p + geom_point() + geom_path(aes(group = 1))
p + geom_raster(aes(fill = z))
```


```{r}
ggplot(df, aes(y, y)) + 
  geom_point(size = 4, colour = "grey20") +
  geom_point(aes(colour = z), size = 2) 

ggplot(df, aes(y, y)) + 
  geom_point(size = 4, colour = "grey20", show.legend = TRUE) +
  geom_point(aes(colour = z), size = 2) 
```

```{r}
norm <- data.frame(x = rnorm(1000), y = rnorm(1000))
norm$z <- cut(norm$x, 3, labels = c("a", "b", "c"))
ggplot(norm, aes(x, y)) + geom_point(aes(colour = z), alpha = 0.1)

ggplot(norm, aes(x, y)) + 
  geom_point(aes(colour = z), alpha = 0.1) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

```{r}
ggplot(df, aes(x, y)) + geom_point(aes(colour = z))
ggplot(df, aes(x, y)) + geom_point(aes(shape = z))
ggplot(df, aes(x, y)) + geom_point(aes(shape = z, colour = z))

```

A number of settings that affect the overall display of the legends are controlled through the theme system

```{r}
df <- data.frame(x = 1:3, y = 1:3, z = c("a", "b", "c"))
base <- ggplot(df, aes(x, y)) + 
  geom_point(aes(colour = z), size = 3) + 
  xlab(NULL) + 
  ylab(NULL)

base + theme(legend.position = "right") # the default 
base + theme(legend.position = "bottom")
base + theme(legend.position = "none")
```

```{r}
base <- ggplot(df, aes(x, y)) + 
  geom_point(aes(colour = z), size = 3)

base + theme(legend.position = c(0, 1), legend.justification = c(0, 1))
base + theme(legend.position = c(0.5, 0.5), legend.justification = c(0.5, 0.5))
base + theme(legend.position = c(1, 0), legend.justification = c(1, 0))
```

## Guide functions

```{r}
df <- data.frame(x = 1, y = 1:3, z = 1:3)
base <- ggplot(df, aes(x, y)) + geom_raster(aes(fill = z))
base 
base + scale_fill_continuous(guide = guide_legend())
base + guides(fill = guide_legend())
```

### Guide Legend

```{r}
df <- data.frame(x = 1, y = 1:4, z = letters[1:4])
# Base plot
p <- ggplot(df, aes(x, y)) + geom_raster(aes(fill = z))
p
p + guides(fill = guide_legend(ncol = 2))
p + guides(fill = guide_legend(ncol = 2, byrow = TRUE))
```

```{r}
p <- ggplot(df, aes(1, y)) + geom_bar(stat = "identity", aes(fill = z))
p
p + guides(fill = guide_legend(reverse = TRUE))
```

```{r}
df <- data.frame(x = 1, y = 1:4, z = 4:1)
p <- ggplot(df, aes(x, y)) + geom_tile(aes(fill = z))

p
p + guides(fill = guide_colorbar(reverse = TRUE))
p + guides(fill = guide_colorbar(barheight = unit(4, "cm")))
```


## Exercises 6.4.4

#### Q1 How do you make legends appear to the left of the plot?


```{r}
df <- data.frame(x = 1:3, y = 1:3, z = c("a", "b", "c"))
base <- ggplot(df, aes(x, y)) + 
  geom_point(aes(colour = z), size = 3) + 
  xlab(NULL) + 
  ylab(NULL)

base + theme(legend.position = "left") 
```

### Q2 

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = drv, shape = drv))
```

### Q3 ???

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point(show.legend = FALSE) + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme(legend.position = "bottom") + 
  guides(colour = guide_legend(nrow = 1))
```

## Limits

The limits, or domain, of a scale are usually derived from the range of the data


```{r}
df <- data.frame(x = 1:3, y = 1:3)
base <- ggplot(df, aes(x, y)) + geom_point() 

base
base + scale_x_continuous(limits = c(1.5, 2.5))
base + scale_x_continuous(limits = c(0, 4))

base + xlim(0, 4)
base + xlim(4, 0) # reversed
base + lims(x = c(0, 4))
```

The axes actually extends a little bit past the limits that you've specified. This ensures that the data does not overlap the axes

```{r}
ggplot(faithfuld, aes(waiting, eruptions)) + 
  geom_raster(aes(fill = density)) + 
  theme(legend.position = "none") 

ggplot(faithfuld, aes(waiting, eruptions)) + 
  geom_raster(aes(fill = density)) + 
  scale_x_continuous(expand = c(0,0)) + 
  scale_y_continuous(expand = c(0,0)) +
  theme(legend.position = "none")
```


By default, any data outside the limits is converted to NA. his means that setting the limits is not the same as visually zooming in to a region of the plot.

```{r}
df <- data.frame(x = 1:5)
p <- ggplot(df, aes(x, 1)) + geom_tile(aes(fill = x), colour = "white")
p
p + scale_fill_gradient(limits = c(2, 4))
p + scale_fill_gradient(limits = c(2, 4), oob = scales::squish)
```



# Exercises 6.5.1

```{r}

mpg$class <-  as.factor(mpg$class)
levels(mpg$class)

fwd <- subset(mpg, drv == "f")
rwd <- subset(mpg, drv == "r")

x_limits <- c(min(mpg$displ), max(mpg$displ))
y_limits <- c(min(mpg$hwy), max(mpg$hwy))

dosomething <-  function(data_set) {
   return(ggplot(data_set, aes(displ, hwy, colour = class)) + 
    geom_point() +
    scale_x_continuous(limits = x_limits) +
    scale_y_continuous(limits = y_limits) +
    lims(color = levels(mpg$class)))
}

dosomething(fwd)
dosomething(rwd)

```

### Question 2 - What does expand_limits() do and how does it work? 

Sometimes you may want to ensure limits include a single value, for all panels or all plots. This function is a thin wrapper around geom_blank that makes it easy to add such values.


Works by adding a blank geom with new data scales

```{r}
expand_limits <- function(...) {
  data <- data.frame(..., stringsAsFactors = FALSE)

  geom_blank(aes_all(names(data)), data, inherit.aes = FALSE)
}
```



```{r}
p <- ggplot(mtcars, aes(mpg, wt)) + geom_point()
p + expand_limits(x = 0)
p + expand_limits(y = c(1, 9))
p + expand_limits(x = 0, y = 0)

ggplot(mtcars, aes(mpg, wt)) +
  geom_point(aes(colour = cyl)) +
  expand_limits(colour = seq(2, 10, by = 2))

ggplot(mtcars, aes(mpg, wt)) +
  geom_point(aes(colour = factor(cyl))) +
  expand_limits(colour = factor(seq(2, 10, by = 2)))
```

### Question 3 - What happens if you add two xlim() calls to the same plot? Why?

it replaces it

```{r}
df <- data.frame(x = 1:3, y = 1:3)
base <- ggplot(df, aes(x, y)) + geom_point() 

base + xlim(1, 10) 
base + xlim(2, 3) + xlim(2,9)
```

### Question 4 - What does scale_x_continuous(limits = c(NA, NA)) do?

uses min and max ? 

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point()

ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point() +
  scale_x_continuous(limits = c(NA, NA))

summary(mpg$displ)
```

## Continuous position scales


apply transformations


if you transform the data, the axes will be labelled in the transformed space. The transformation occurs before any statistical summaries


```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  scale_y_continuous(trans = "reciprocal")


ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  scale_y_continuous(trans = "sqrt")

# Log transform x and y axes
ggplot(diamonds, aes(price, carat)) + 
  geom_bin2d() + 
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")
```


```{r}
base <- ggplot(economics, aes(date, psavert)) + 
  geom_line(na.rm = TRUE) +
  labs(x = NULL, y = NULL)

base # Default breaks and labels
base + scale_x_date(date_labels = "%y", date_breaks = "5 years")
base + scale_x_date(date_labels = "%Y", date_breaks = "20 years")

base + scale_x_date(
  limits = as.Date(c("2004-01-01", "2005-01-01")),
  date_labels = "%b %y",
  date_minor_breaks = "1 month"
)
base + scale_x_date(
  limits = as.Date(c("2004-01-01", "2004-06-01")),
  date_labels = "%m/%d",
  date_minor_breaks = "2 weeks"
)
```


## Color

There are quite a few different ways of mapping values to colours in ggplot2: four different gradient-based methods for continuous values, and two methods for mapping discrete values.


### Colour theory 

At the physical level, colour is produced by a mixture of wavelengths of light.

To characterise a colour completely, we need to know the complete mixture of wavelengths.

```{r}
erupt <- ggplot(faithfuld, aes(waiting, eruptions, fill = density)) +
  geom_raster() +
  scale_x_continuous(NULL, expand = c(0, 0)) + 
  scale_y_continuous(NULL, expand = c(0, 0)) + 
  theme(legend.position = "none")

erupt

erupt + scale_fill_gradient(low = "white", high = "black")

erupt + scale_fill_gradient(
  low = munsell::mnsl("5G 9/2"), 
  high = munsell::mnsl("5G 6/8")
)

mid <- median(faithfuld$density)
erupt + scale_fill_gradient2(midpoint = mid) 
```
```{r}
summary(luv_colours)
head(luv_colours)

ggplot(luv_colours, aes(u, v, color = col)) +
  geom_point(data = colors())

```

