---
title: "3_2 Basic Plots"
author: "Wal McConnell"
date: "18 June 2017"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
df <- data.frame(
  x = c(3,1,5),
  y = c(2,4,6),
  label = c("a", "b", "c")
)
p <- ggplot(df, aes(x,y, label = label)) +
  labs(x=NULL, y = NULL) + # Hide Axis Labels
  theme(plot.title = element_text(size = 20))
```

## 3.2 Basic Plot Types

```{r}
plot(p + geom_point() + ggtitle('Point'))
```
```{r}
plot(p + geom_text() + ggtitle('Text'))
```

```{r}
plot(p + geom_bar(stat = "identity") + ggtitle('Bar'))
```

```{r}
plot(p + geom_tile() + ggtitle('Raster'))
```
### What does Raster mean ? 

```{r}
plot(p + geom_line() + ggtitle('Line'))
```

```{r}
plot(p + geom_area() + ggtitle('Area'))
```

```{r}
plot(p + geom_path() + ggtitle('Path'))
```

```{r}
plot(p + geom_polygon() + ggtitle('Polygon?'))
```

## Exercises 3.2.1

### Question 1 - Chart Types
#### Pie Chart?
```{r}
plot(p + geom_bar(width = 1, stat = "identity") + ggtitle('Bar'))
```

```{r}
p <- ggplot(df, aes(x = "",y,label = label)) +
  labs(x=NULL, y = NULL) + # Hide Axis Labels
  theme(plot.title = element_text(size = 20))
plot(p + geom_bar(width = 1, stat = "identity") + coord_polar("x", start = 0) + ggtitle('Pie'))
```


### Question 2 geom_path v geom_polygon

geom_path connects the observations in the order in which they appear in the data. 
geom_line connects them in order of the variable on the x axis. 
geom_step() creates a stairstep plot, highlighting exactly when changes occur. The group aesthetic determines which cases are connected together.


Polygons are very similar to paths (as drawn by geom_path) except that the start and end points are connected and the inside is coloured by fill. The group aesthetic determines which cases are connected together into a polygon.

### Question 3

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + geom_smooth() # method = loess
```


geom_smooth = geom_area + geom_line ? 
geom_boxplot = 

