---
title: "Movies"
author: "Wal McConnell"
date: "24 June 2017"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggplot2movies)
library(precis)
knitr::opts_chunk$set(echo = TRUE)
```

## Movies dataset

A data frame with 28819 rows and 24 variables

* title. Title of the movie.
* year. Year of release.
* budget. Total budget (if known) in US dollars
* length. Length in minutes.
* rating. Average IMDB user rating.
* votes. Number of IMDB users who rated this movie.
* r1-10. Multiplying by ten gives percentile (to nearest 10%) of users who rated this movie a 1.
* mpaa. MPAA rating.
* action, animation, comedy, drama, documentary, romance, short. Binary variables representing if movie was classified as belonging to that genre.

Movies were selected for inclusion if they had a known length and had been rated by at least one imdb user.


```{r}
precis(movies)
```


```{r}
head(movies)
```

Prepare per type filtered data sets

```{r}
action_movies <- movies %>% filter(Action == 1)
comedy_movies <- movies %>% filter(Comedy == 1)
drama_movies <- movies %>% filter(Drama == 1)
documentary_movies <- movies %>% filter(Documentary == 1)
romance_movies <- movies %>% filter(Romance == 1)
short_movies <- movies %>% filter(Short == 1)
```

## Questions

### What is the growth in the number of movies over time ?

```{r}
movie_count_by_year <- function(movie_data) {
  summarized_data <- movie_data %>% group_by(year) %>% 
    summarize(movies_count = sum(n())) %>% 
    select(year, movies_count)
  
  return(summarized_data)
}

movies_by_year <- movie_count_by_year(movies)

ggplot(movies_by_year, aes(year, movies_count)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  labs(x = "Year", y = "Number of Movies")

action_movies_by_year <- movie_count_by_year(action_movies)
comedy_movies_by_year <- movie_count_by_year(comedy_movies)
drama_movies_by_year <- movie_count_by_year(drama_movies)
documentary_movies_by_year <- movie_count_by_year(documentary_movies)
romance_movies_by_year <- movie_count_by_year(romance_movies)
short_movies_by_year <- movie_count_by_year(short_movies)

ggplot() + 
  geom_smooth(data = action_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Action")) + 
  geom_smooth(data = comedy_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Comedy")) + 
  geom_smooth(data = drama_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Drama")) + 
  geom_smooth(data = documentary_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Documentary")) + 
  geom_smooth(data = romance_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Romance")) + 
  geom_smooth(data = romance_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Short")) + 
  geom_smooth(data = movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Total")) + 

  labs(x = "Year", y = "Number of Movies") +
  ggtitle("movies per year")


```


### Are movie titles getting longer ?

```{r}
library(stringr)
movie_title_summary <- movies %>% 
  group_by(year) %>% 
  summarise(mean_title_length = mean(str_length(title)), max_title_length = max(str_length(title)))

ggplot(movie_title_summary) + 
  geom_smooth(aes(year, mean_title_length, color = "Mean title length")) +
  geom_smooth(aes(year, max_title_length, color = "Max title length")) +
  labs(x = "Year" , y = "Title Length") +
  ggtitle("Movie title length over time ")
```

* Are movies getting longer ?
```{r}
movie_length_summary <- movies %>% 
  group_by(year) %>% 
  summarise(mean_length = mean(length), max_length = max(length))

ggplot(movie_length_summary) + 
  geom_smooth(aes(year, mean_length, color = "Mean length")) +
  geom_smooth(aes(year, max_length, color = "Max length")) +
  labs(x = "Year" , y = "Length") +
  ggtitle("Movie length over time ")
```


### Are movies getting more expensive ?
```{r}
mean(movies$budget, na.rm = TRUE)


movie_budget_summary <- movies %>% 
  group_by(year) %>% 
  summarise(mean_budget = mean(budget, na.rm = TRUE), 
            max_budget = max(budget, na.rm = TRUE),
            min_budget = min(budget, na.rm = TRUE),
            n90th_percentile_budget = quantile(budget, 0.9, na.rm = TRUE)
            ) %>% 
  filter(!is.nan(mean_budget), mean_budget > 0)

ggplot(movie_budget_summary) + 
  geom_smooth(aes(year, min_budget, color = "Min")) +
  geom_smooth(aes(year, mean_budget, color = "Mean")) +
  geom_smooth(aes(year, n90th_percentile_budget, color = "90th Percentile")) +
  geom_smooth(aes(year, max_budget, color = "Max")) +
  labs(x = "Year" , y = "Budget") +
  ggtitle("Movie budget over time ")
```

## What is the relationship between movie length and votes ?

```{r}
ggplot(movies, aes(length, votes)) + geom_point()
```



* How does MPAA influence votes ?

```{r}
movies_by_rating_and_votes <- movies %>%
  filter(!is.null(mpaa), mpaa != '', votes > 0) %>%
  group_by(mpaa, votes)  %>%
  select(title, year, mpaa, votes)
  

ggplot(movies_by_rating_and_votes, aes(mpaa, votes)) + geom_boxplot()
```



