---
title: "Movies"
author: "Wal McConnell"
date: "24 June 2017"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggplot2movies)
library(precis)
knitr::opts_chunk$set(echo = TRUE)
```

## Movies dataset

A data frame with 28819 rows and 24 variables

* title. Title of the movie.
* year. Year of release.
* budget. Total budget (if known) in US dollars
* length. Length in minutes.
* rating. Average IMDB user rating.
* votes. Number of IMDB users who rated this movie.
* r1-10. Multiplying by ten gives percentile (to nearest 10%) of users who rated this movie a 1.
* mpaa. MPAA rating.
* action, animation, comedy, drama, documentary, romance, short. Binary variables representing if movie was classified as belonging to that genre.

Movies were selected for inclusion if they had a known length and had been rated by at least one imdb user.


```{r}
precis(movies)
```


```{r}
head(movies)
```

Prepare filtered data sets

```{r}
action_movies <- movies %>% filter(Action == 1)
comedy_movies <- movies %>% filter(Comedy == 1)
drama_movies <- movies %>% filter(Drama == 1)
documentary_movies <- movies %>% filter(Documentary == 1)
romance_movies <- movies %>% filter(Romance == 1)
short_movies <- movies %>% filter(Short == 1)

movies_with_budget <- movies %>% filter(budget > 0)
less_than_five_hour_movies <- movies %>% filter(length < (5 * 60))
movies_without_outliers <- movies %>% filter(budget < 500000000, length < (5 * 60))

```


Can movies have more than one classification ?

```{r}
classification_totals <- movies %>% 
  mutate(classification_total = (Action + Comedy + Drama + Documentary + Romance + Short)) %>%
  select(classification_total)
table(classification_totals)
```



## Distributions

### Movie Length
```{r}
ggplot(movies, aes(length)) + 
  geom_histogram(binwidth = 5) + 
  xlim(0,300) +
  ggtitle("Movies < 5 hours")
```

### Movie Budget

```{r}
ggplot(movies_with_budget, aes(budget)) + 
  geom_histogram(binwidth = 100000) +
  xlim(0,100000000) +
  ggtitle("Movies < 5 hours")
```



## Questions

### How has the number of movies released each year changed over time ?

```{r}
movie_count_by_year <- function(movie_data) {
  summarized_data <- movie_data %>% group_by(year) %>% 
    summarize(movies_count = sum(n())) %>% 
    select(year, movies_count)
  
  return(summarized_data)
}

movies_by_year <- movie_count_by_year(movies)

ggplot(movies_by_year, aes(year, movies_count)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  labs(x = "Year", y = "Number of Movies") +
  ggtitle("Movies per year")

action_movies_by_year <- movie_count_by_year(action_movies)
comedy_movies_by_year <- movie_count_by_year(comedy_movies)
drama_movies_by_year <- movie_count_by_year(drama_movies)
documentary_movies_by_year <- movie_count_by_year(documentary_movies)
romance_movies_by_year <- movie_count_by_year(romance_movies)
short_movies_by_year <- movie_count_by_year(short_movies)

ggplot() + 
  geom_smooth(data = action_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Action")) + 
  geom_smooth(data = comedy_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Comedy")) + 
  geom_smooth(data = drama_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Drama")) + 
  geom_smooth(data = documentary_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Documentary")) + 
  geom_smooth(data = romance_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Romance")) + 
  geom_smooth(data = romance_movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Short")) + 
  geom_smooth(data = movies_by_year, se = FALSE, aes(x = year, y = movies_count, color = "Total")) + 

  labs(x = "Year", y = "Number of Movies") +
  ggtitle("Movies per year by Type")

```





## Movie Title Length


### What is the distribution of movie title length

```{r}
library(stringr)
ggplot(movies, aes("Title", str_length(title))) + geom_boxplot() + ggtitle("Movie Title Length (chars)")
ggplot(movies, aes(str_length(title))) + 
  geom_histogram(binwidth=2) + 
  geom_freqpoly(binwidth=2, color = "blue") + 
  labs(x = 'Characters', 'Count') +
  ggtitle("Movie Title Length")
ggplot(movies, aes(str_length(title))) + 
  geom_density(color = 'hotpink') +
  labs(x = 'Characters', 'Count') +
  ggtitle("Movie Title Length Density")
```

### How does mpaa affect title length ?
```{r}
movies_with_mpaa <- movies %>% filter(!is.null(mpaa), mpaa != '')
ggplot(movies_with_mpaa, aes(mpaa, str_length(title))) + 
  geom_boxplot() + 
  labs(x = "MPAA", y = "Movie Title Length") +
  ggtitle("Movie Title Length v MPAA Rating")

ggplot(movies_with_mpaa, aes(str_length(title), color = mpaa)) +
  geom_density(binwidth = 2) +
  labs(x = 'Movie Title Length', y = 'Movie Count')
```


### Are movie titles getting longer ?


```{r}
library(stringr)
movie_title_summary <- movies %>% 
  group_by(year) %>% 
  summarise(mean_title_length = mean(str_length(title)), max_title_length = max(str_length(title)))

ggplot(movie_title_summary) + 
  geom_smooth(aes(year, mean_title_length, color = "Mean title length")) +
  geom_smooth(aes(year, max_title_length, color = "Max title length")) +
  labs(x = "Year" , y = "Title Length") +
  ggtitle("Movie title length over time ")
```

* Are movies getting longer ?
```{r}
movie_length_summary <- movies %>% 
  group_by(year) %>% 
  summarise(mean_length = mean(length), max_length = max(length))

ggplot(movie_length_summary) + 
  geom_smooth(aes(year, mean_length, color = "Mean length")) +
  geom_smooth(aes(year, max_length, color = "Max length")) +
  labs(x = "Year" , y = "Length") +
  ggtitle("Movie length over time ")
```


### Are movies getting more expensive ?
```{r}
mean(movies$budget, na.rm = TRUE)

movie_budget_summary <- movies %>% 
  group_by(year) %>% 
  summarise(mean_budget = mean(budget, na.rm = TRUE), 
            max_budget = max(budget, na.rm = TRUE),
            min_budget = min(budget, na.rm = TRUE),
            n90th_percentile_budget = quantile(budget, 0.9, na.rm = TRUE)
            ) %>% 
  filter(!is.nan(mean_budget), mean_budget > 0)

ggplot(movie_budget_summary) + 
  geom_smooth(aes(year, min_budget, color = "Min")) +
  geom_smooth(aes(year, mean_budget, color = "Mean")) +
  geom_smooth(aes(year, n90th_percentile_budget, color = "90th Percentile")) +
  geom_smooth(aes(year, max_budget, color = "Max")) +
  labs(x = "Year" , y = "Budget") +
  ggtitle("Movie budget over time ")
```


### How has budget affected votes by decade 
```{r}

movies_with_decade <- movies_with_budget %>% mutate(decade = signif(year,3))
ggplot(movies_with_decade, aes(votes, (budget / 1000000))) + 
  geom_point() + 
  facet_wrap(~ decade) +
  labs(x = "Votes", y = "Million /$") +
  ggtitle("Votes v Budget")
```



### Is it more expensive to make a longer movie ?

```{r}
ggplot(less_than_five_hour_movies, aes(length, budget)) + 
  geom_point() +
  geom_smooth() +
  labs(x = "Minutes", y = "Budget") +
  ggtitle("(Movies < 5Hrs) - Movie Length v Budget")

```


## Votes

### Does Movie Length affect votes ?
```{r}

ggplot(less_than_five_hour_movies, aes(length, votes)) + 
  geom_point() + 
  geom_smooth()
```



### How does MPAA influence votes ?

```{r}
movies_by_rating_and_votes <- movies %>%
  filter(!is.null(mpaa), mpaa != '', votes > 0) %>%
  group_by(mpaa, votes)  %>%
  select(title, year, mpaa, votes)
  

ggplot(movies_by_rating_and_votes, aes(mpaa, votes)) + geom_boxplot()
```

### GGaly

```{r}
library(GGally)
ggpairs(select(movies_without_outliers, budget, length, votes, rating, mpaa))
ggpairs(select(movies_without_outliers, budget, length, votes, rating, mpaa), aes(colour = mpaa, alpha = 0.4))
```

